---
title: "Chapter 2&3 Practice and homeworks"
output:
  html_document:
    df_print: paged
---
***
The layout for the following notebooks will first go through the book chapter's practice section and then tackle the homeworks provided online. In between, I might copy code sections from the book to facilitate the practical aspects associated to the learning and application of the methods of interest.

*Disclaimer: The answers provided here are not only mine and are not necessarily correct* 

See Sources here containing some of the copied answers in this notebook: 

- [Source 1](https://rpubs.com/andersgs/my_solutions_chapter2_statrethink)
- [Source 2](http://rstudio-pubs-static.s3.amazonaws.com/326789_e280ba45916a400f9556212409406da2.html)
    
### Practice 2.6.

##### **Easy:**

* **2E1.** The expressions that correspond to the "The probability of rain on Monday" are: Pr(rain|Monday) and Pr(rain,Monday) / Pr(Monday) --

* **2E2.** The probability that it is Monday given that it is raining corresponds to the expression Pr(Monday|Rain)

* **2E3.** Pr(Monday|Rain) and Pr(rain|Monday) * Pr(Monday) / Pr(rain) [according to Bayes rule] corresponds to the statement: "The probability that it is Monday given that it is raining"

* **2E4.** With regards to the global tossing example from the chapter, saying that "the probability of water is 0.7 " implies that given our limited knowledge and based on the outcomes of the experiment, our best guess of the true proportion of water in the globe is 0.7. --

***

##### **Medium:**

> Recalling the global tossing model from chapter 2

```{r echo=FALSE}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )

# compute likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid 
                      )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# Plot
plot( p_grid , posterior , type="b" ,
xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )

```

***

* **2M1.& 2M2**


    + In case the set of observation was: (1) W, W, W
```{r echo=FALSE}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )
prior <- ifelse(p_grid < 0.5, 0, 1)

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid 
                      )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# Plot
plot( p_grid , posterior , type="b" ,
xlab="probability of water" , ylab="posterior probability" )
mtext( "1-W, W, W (20 points)" )

```

     + In case the set of observation was: (2) W, W, W, L
```{r echo=FALSE}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )
prior <- ifelse(p_grid < 0.5, 0, 1)

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=4 , prob=p_grid 
                      )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# Plot
plot( p_grid , posterior , type="b" ,
xlab="probability of water" , ylab="posterior probability" )
mtext( "2-W, W, W, L (20 points)" )

```

    + In case the set of observation was: (3) L, W, W, L, W, W, W
```{r echo=FALSE}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )

# define prior
prior <- rep( 1 , 20 )
prior <- ifelse(p_grid < 0.5, 0, 1) 

# compute likelihood at each value in grid
likelihood <- dbinom( 5 , size=7 , prob=p_grid 
                      )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

# Plot
plot( p_grid , posterior , type="b" ,
xlab="probability of water" , ylab="posterior probability" )
mtext( "3-L, W, W, L, W, W, W (20 points)" )

```

* **2M3** --
```{r echo=TRUE}
p_E = 1/2 # probability that globe is Earth
p_M = 1/2 # probability that globe is Mars

p_land_E = 0.3 # probability land given Earth p(land|Earth)
p_land_M = 1.0 # probability land given Mars p(land|Mars)

p_land = p_land_E*p_E + p_land_M*p_M # probability of land 

# What is p(Earth|land) ?
## p(Earth|land) = (p(land|Earth) * p(Earth)) / p(land)
p_earth_L = (p_land_E * p_E)/p_land
p_earth_L

```

* **2M4** 
Card1: B/B Card2: B/W Card3: W/W

If we observed one black side up, there are 2 ways of producing the black side up if we pick the B/B card and one way if we pick the B/W card, so in total, 2 of 3 ways are consistent with the other side of the card being black. The answer is 2/3.

* **2M5** 
With an extra B/B card:

If we observe one black side up, there are 4 ways of producing the black side up (2 for each B/B card) and 1 for picking the B/W card. The total probability that the other side is black is thus: 4/5

* **2M6** 
Card1: B/B Card2: B/W Card3: W/W

Black sides are heavier than cards with white sides so it is less likely that a card with black sides is pulled from the bag.

There are still 2 ways for B/B to produce a black side up, 1 way for B/W and zero ways for W/W. But according to the prior information, there is 1 way to get the B/B card, 2 ways to get the B/W card and 3 ways to get the W/W card so:

The total ways for the B/B card to produce a black side up are 2 * 1 and the total ways for the B/W card to produce a black side up are 1 * 2.

So, there are 4 ways total to see a black side up and 2 of these come from the B/B card hence the probability that the other side is black is now at 0.5 

* **2M7**  --
Card1: B/B Card2: B/W Card3: W/W

The B/B card has 2 ways to produce an observation with the black side up. This leaves the B/W and the W/W card with the job of producing the next observation. The B/W has only 1 way to produce a white side up while the W/W has 2 ways. So there are 3 ways to get the second card to show the white side up. Assuming the first card is B/B, there are 6 ways to see the BW sequence.

If however the B/W card is drawn first, there is 1 way for it to show the black side up. This leaves the other two cards to show the white side up. The B/B can’t show white while W/W has 2 ways to show white up so that is 2 ways to see the sequence Black and White when the first card drawn is B/W.

The probability is thus: 8 total ways of producing the sequence BW with 6 ways with B/B being drawn first.
